---
title: "GAPS 3.0 Chinook Baseline - rubias"
author: "Kyle Shedd"
date: "April 6, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = "V:/Analysis/1_SEAK/Chinook/Baseline/GAPS3.0")
```

## Objectives

Here I am testing Eric Anderson's new R package [rubias](https://github.com/eriqande/rubias#rubias-----genetic-stock-identification-gsi-in-the-tidyverse) on the GAPS 3.0 coastwide Chinook microsatellite baseline used in for Southeast Alaska fisheries for the Pacific Salmon Treaty. I will convert a `Genepop` file to our `.gcl` objects, then show how to convert our `.gcl` objects to `rubias` format, and demonstrate some of the built-in baseline testing capabilities.

## Convert .gcl baseline to rubias

First load `GCL` functions.

```{r load gcl functions, results='hide'}
source("C:/Users/krshedd/Documents/R/Functions.GCL.R")
```

Next load libraries, `LocusControl` and baseline objects (loci, groupvec, etc.).

```{r load baseline objects}
library(tidyverse)
library(rubias)

setwd("V:/Analysis/1_SEAK/Chinook/Mixture/SEAK17/Objects")

LocusControl <- dget(file = "LocusControl.txt")
GAPSLoci <- dget(file = "GAPSLoci.txt")
GAPSLoci_reordered <- dget(file = "GAPSLoci_reordered.txt")
SEAKPops357 <- dget(file = "SEAKPops357.txt")
GroupVec33RG_357 <- dget(file = "GroupVec33RG_357.txt")
GroupNames33 <- dget(file = "GroupNames33.txt")
# Colors15 <- dget(file = "Objects/Colors15.txt")
```

Unfortunately, `GenepopTOgcl.GCL` does not work for whatever reason. I will have to load the Genepop file manually to convert directly to `rubias` format.

```{r import genepop}
GAPSxGOA13loci_2.gen <- scan(file = "Associated Data/GAPS3xGOA13loci_2.gen", what = "", sep = "\n")
pop.index <- grep(pattern = "Pop", x = GAPSxGOA13loci_2.gen, ignore.case = FALSE)
all.equal(GAPSxGOA13loci_2.gen[2:(pop.index[1] - 1)], GAPSLoci)  # verify GAPSLoci is correct order of Genepop file
genotypes.chr <- GAPSxGOA13loci_2.gen[-c(1:(pop.index[1] - 1), pop.index)]  # get rid of header, loci, and "Pop" rows
genotypes.chr_split1 <- str_split(string = genotypes.chr, pattern = "  ,   ", simplify = TRUE)  # split to get indiv and genotypes in two separate columns
collection <- str_split(string = genotypes.chr_split1[, 1], pattern = "\\.", simplify = TRUE)[, 1]  # split to get collection column
indiv <- genotypes.chr_split1[, 1]  # indiv column
genotypes.mat <- str_split(string = genotypes.chr_split1[, 2], pattern = "  ", simplify = TRUE)

str(apply(genotypes.mat, 2, function(locus) {cbind(str_sub(string = locus, start = 1, end = 3),
                                                   str_sub(string = locus, start = 4, end = 6))}))

str(str_sub(string = genotypes.mat, start = 1, end = 3))

```


Next load the function `create_rubias_baseline` to create a tidy dataframe of the baseline in [`rubias` format](https://github.com/eriqande/rubias#input-data). One individual per row, with the first four columns required, then genotypes in a two column format. **Note** that the `rubias` specific GCL functions are currently only on the [develop](https://github.com/krshedd/GCL-R-Scripts/tree/develop) branch of the GCL-R-Scripts repository on <github.com>.


```{r convert baseline to rubias}
source("C:/Users/krshedd/Documents/R/GCL-R-Scripts/create_rubias_baseline.R")
kma_473pops_89loci_15groups.rubias_base <- create_rubias_baseline(sillyvec = KMA473Pops, loci = loci89, group_names = Groups15, groupvec = KMA473PopsGroupVec15)
str(kma_473pops_89loci_15groups.rubias_base, max.level = 0)
head(kma_473pops_89loci_15groups.rubias_base[, 1:10])
```